### 2015-07-20 text/ms.Rmd
Jess edits - minor typos and many comments at top of doc

@@

In fisheries science, the commonly used operational model for determining status and trajectory of exploited fish populations are stock assessments that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$). However, a broad range of data are not available for the majority of fish populations [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: catches and a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. The four models frequently disagreed about population status and any one model had relatively poor predictive ability on average (e.g. Fig. \ref{motivate}).

Here, we estimate population status and trajectory of exploited fish populations using ensembles and superensembles (collectively 'ensemble methods') of these four data-limited models. We explore a variety of approaches applied to both simulated and real-world fish stocks and compare predictive performance. We find that ensembles, and in particular superensembles, generally improved predictive performance of status and trajectory of fish populations over any single model---especially when considered across multiple dimensions of predictive ability. Furthermore, the output from superensembles can aid mechanistic understanding of how individual models of status and trajectory perform.

@@

We first developed and tested ensemble methods with a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulations included three fish life histories: small pelagic, demersal, and large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamics: a constant harvest rate, a harvest rate coupled with biomass to mimic an open-access single-species fishery, a scenario where harvest rate increases continuously, and a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation included a scenario without observation error and with multiplicative observation error around catch at $N(0, 0.2)$.

@@

* *CMSY* (catch-MSY) implements a stock-reduction analysis with Schaefer biomass dynamics [@martell2013]. It requires a prior distribution for $r$ and assigns a prior to the relative proportion of biomass at the end compared to unfished biomass (depletion) based on the percentage of maximum catch at the end of the time series. Our version of the model was modified from @martell2013 to generate biomass trends from all viable $r$-$K$ pairs and produce an estimate of $B/B_\mathrm{MSY}$ from the median trend.

@@

* *mPRM* (modified panel regression model) is a modified version of the panel-regression model from @costello2012. Unlike the other models, mPRM is empirical and not mechanistic---it uses the RAM Legacy Stock Assessment database to fit a regression model to a series of characteristics of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as the response. The model used in this paper is modified from the original in that it condenses the life-history categories into three categories to match the simulated dataset, removes the maximum catch predictor since the absolute catch in the simulated dataset is arbitrary, and does not implement the bias correction needed in @costello2012 for deriving aggregate estimates of median status across multiple stocks.

@@

\noindent 
where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{CMSY}$ is the mean or slope prediction from the CMSY model etc.\ (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the spectral density of the scaled catch series at frequencies of 0.20 and 0.05, $b$ is the mean or slope value the model is fitted to, and $\theta$ is the set of parameters for the regression function. For all superensemble models of mean $B/B_\mathrm{MSY}$---a ratio bounded at zero---we fit the models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.

We fit the linear model superensemble with second-order interactions. We fit two machine learning superensemble models based on regression trees. Regression trees sequentially determine what value of a predictor best splits the response data into two 'branches' based on a specified loss function [@breiman1984]. In random forests, a series of regression trees are built on a random subset of the data and random subset of the covariates of the model [@breiman2001]. In generalized boosted models (GBMs), each subsequent model is fit to the residuals from the previous model; data points that are fit poorly in a given model are given more weight in the next model [@elith2008]. Random forests and GBMs can provide strong predictive performance and fit highly non-linear relationships [@elith2008; @hastie2009]. We fit random forest models with the `randomForest` package [@liaw2002] for \textsf{R} with the default argument values. We fit boosted regression tree models with the `gbm` package [@ridgeway2015] for \textsf{R}. Based on cross-validation with the **caret** \textsf{R} package, we fit GBMs with 2000 trees, an interaction depth of $6$, a learning rate (shrinkage parameter) of $0.01$, and all other arguments at their default values.

@@

Applied to the simulated dataset of known status, the individual models had variable success at recovering the mean (status) and slope (trajectory) of $B/B_\mathrm{MSY}$ in the last five years. All models exhibited a high degree of scatter around the one-to-one line of perfect status prediction (Fig. \ref{hexagon}). CMSY exhibited bimodal predictions of status but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a). COM-SIR and SSCOM both correctly identified a number of low status stocks, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). There was generally little correlation between true and predicted trajectory for any of the individual models (rank-order correlation = `r range_slope_corr_non_sscom`) besides SSCOM (correlation = `r slope_corr_sscom`; Figs \ref{scatter-sim-slope}a--d).

@@

When applied to the stock assessment database, the superensemble models---trained exclusively on the simulated dataset---generally performed as well or better than the best individual models. The mean, random forest, and GBM ensembles outperformed the mPRM method which is trained directly on the RAM Legacy Stock Assessment database itself (Fig. \ref{performance}b, Fig. \ref{hexagon-ram}). Compared to the individual models, the machine learning superensembles increased accuracy by `r mean_ram$mach_vs_ind_mare_fold`%, improved correlation from `r mean_ram$ind_corr_range` to `r mean_ram$mach_corr_range`, and reduced bias from `r mean_ram$ind_mre_range` to `r mean_ram$mach_mre_range`.

@@

Ensemble methods provide a useful approach to situations where conservation and ecological resource management decisions must be made on the basis of multiple, potentially contrasting estimates of status---a situation common to many settings beyond fisheries. Compared to individual models of fish population status, ensemble methods were consistently the best or among the best across three performance dimensions (accuracy, bias, and rank-order correlation), two response variables (status and trajectory), two datasets (simulated and global fisheries), and multiple ensemble methods (from a simple average to machine learning superensembles). Our results suggest choosing an superensemble model that allows for non-parametric relationships, such as machine learning methods. These methods provide added insight into individual model behaviour and generally performed the best; however, even a simple average of predictions across multiple models may provide a more useful measure of the desired parameter than a single model in many cases.

@@

Combining predictions from multiple models via superensemble methods is also broadly useful in other subfields of ecology and fisheries science. Predictions about extinction risk are widely used at the national (e.g. the US Endangered Species Act and the Canadian Species at Risk Act) and international [e.g. the IUCN Red List, @iucn2015] levels. These risk assessments generally involve fitting regression models to outcomes for individual species along with predictors of species risk [e.g. @anderson2011a; @pinsky2011], or population dynamics models to data for individual species [e.g. @dfo2010]. Both types of models are prone to error caused by model-misspecification and therefore results are sensitive to decisions about model structure [@brooks2015]. Although there are options to account for potential model-misspecification in determination of species risk [e.g. coefficient averaging, @burnham2002; generalized modeling, @yeakel2011; or semi-parametric methods, @thorson2014], ensemble methods are a relatively simple way to combine predictions in a transparent manner. Beyond estimates of status and trajectory, ensemble methods could be used to increase the robustness of spatial predictions when designing networks of protected areas [@rassweiler2014], or to forecast potential spatial shifts in species distribution given climate impacts [@harsch2014].

>>>>

<!--Jessicas Comments: 
1. Massive improvement in abstract and introduction. I really like the strucutre and how you explain the models. makes sense to people who aren't familiar with the problems around datalimited fisheries, and the ideas and need for ensemble models. This is an amazing paper - quite complicated, but well explained to show people why it matters and how they can use it. 
2. If I remember correctly, Cons Biol doesn't like having the text wrapped around between two lines with hyphens as the PDF shows. Not sure how to fix that, but just something to note. 
3. is it normal to describe the results and findings at the end of the introduction? i see it sometimes, but not sure if Cons BIol like it. If not, you could put that final paragraph of introduction at start of discussion.
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
6. Fix up Simulation FLR package
7. Update Version xx downloaded on xx RAM legacy database
8. line 125 in pdf - Ram legacy database - how many stocks didn't converge? readers might want to know how many stocks you threw away.
9. make datalimited package and ensembles code public when published
10. Sorrry - dumb questions, but probably should ask to properly understand equations... line 184 pdf - what is theta-hat? and what do you use it for? should you describe this, or is it obvious to more technically minded folk than me? I don't fully understand what b is. 
11. Also, why do you use 'arg min' and 'a loss function'? Am i right in thinking that the first equation is to find out which superensemble models are best predictors (different theta-hats) and the different 'loss functions' are those used to determine bias and accuracy? Or am I completely wrong? Would it help to explain what arg min and the loss functions are used for (as it took me a long time to figure out!)? you could just say 'for performance metrics, see below'
12. line 202 - is this cross validation of GBM the same as the cross-validation described in the section below about performance metrics?
13. FIgure 1 - I like it as is. 
14. Figure 4 - tweak labels when all else is completed.
15. Line 255 from pdf - CMSY predictions were approximately linear  (Fig S4) - Do you mean Fig S4a? Maybe I don't know how to interpret these graphs - I don't see how you can extract that info from the CMSY graph. I don't really know wht you mean by 'high and low prediction clusters'.
16. line 258-59 - if mPRM and SSCOM are high, then true status tends to be high. Is there another example that shows how covariates are exploited. The example shown is convincing in Fig S5, but not in Fig S6l, because the predicted B/Bmsy is only just above (light pink shading). though ths may not matter. you can ignore this comment if you want.
17. check for consistent US spelling - e.g. behaviour - line 279-280. there might be more that I didn't notice. also - line 290 and 291 - overfitting vs over-fitting. I'm not sure which is correct. 
18. line 314-15 need a reference showing that SSCOM contributed little to status from GBM. e.g. which figure shows this?
19. line 334 - should this be "effort dynamics at low biomasses in COM-SIR", not SSCOM? 
20. I said that i'd help with the last paragraph about other uses for ensemble methos. If there is space within the word limit, could we add just one other potential use? Are the references listed in 350-51 using ensemble models or do they simply provide examples of where ensembles could be used. Just check my addition - and delete it if you aren't happy with it. I don't know how to insert references - i'll leave that up to you. 
21. fix acknowledgements section
22. I didn't check references. 
23. other than that! Awesome! Hope my comments were helpful.

@@

In fisheries science, the commonly used operational model for determining status and trajectory of exploited fish populations are stock assessments that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass of a stock that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$). However, the broad range of data required to conduct stock assessments are not available for the majority of fish populations, including those of conservation concern and of economic interest to fisheries [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: catches and a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. The four models frequently disagreed about population status and any one model had relatively poor predictive ability on average (e.g. Fig. \ref{motivate}).

Here, we estimate population status and trajectory of exploited fish populations using ensembles and superensembles (collectively 'ensemble methods') of these four data-limited models. We apply a variety of approaches to both simulated and real-world fish stocks and compare their predictive performance. We find that ensembles, and in particular superensembles, generally improved predictive performance of status and trajectory of fish populations over any single model---especially when considered across multiple dimensions of predictive ability. Furthermore, the output from superensembles can aid mechanistic understanding of how individual models of status and trajectory perform.

@@

We first developed and tested ensemble methods with a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulations included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamics: a constant harvest rate, a harvest rate coupled with biomass to mimic an open-access single-species fishery, a scenario where harvest rate increases continuously, and a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error.

@@

* *CMSY* (catch-MSY) implements a stock-reduction analysis with Schaefer biomass dynamics [@martell2013]. It requires a prior distribution for $r$ and assigns a prior to the relative proportion of biomass at the end of the time series compared to unfished biomass (depletion) based on the percentage of maximum catch at the end of the time series. Our version of the model was modified from @martell2013 to generate biomass trends from all viable $r$-$K$ pairs and produce an estimate of $B/B_\mathrm{MSY}$ from the median trend.

@@

* *mPRM* (modified panel regression model) is a modified version of the panel-regression model from @costello2012. Unlike the other models, mPRM is empirical and not mechanistic --- it uses the RAM Legacy Stock Assessment database to fit a regression model to a series of characteristics of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as the response. The model used in this paper is modified from the original in that it condenses the life-history categories into three categories to match the simulated dataset, removes the maximum catch predictor since the absolute catch in the simulated dataset is arbitrary, and does not implement the bias correction needed in @costello2012 for deriving aggregate estimates of median status across multiple stocks.

@@

\noindent 
where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{CMSY}$ is the mean or slope prediction from the CMSY model etc.\ (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the spectral density of the scaled catch series at frequencies of 0.20 and 0.05, $b$ is the mean or slope value the model is fitted to, and $\theta$ is the set of parameters for the regression function. For all superensemble models of mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.

We fit the linear model superensemble with second-order interactions. We fit the two machine learning superensemble models based on regression trees. Regression trees sequentially determine what value of a predictor best splits the response data into two 'branches' based on a specified loss function [@breiman1984]. In random forests, a series of regression trees are built on a random subset of the data and random subset of the covariates of the model [@breiman2001]. In generalized boosted models (GBMs), each subsequent model is fit to the residuals from the previous model; data points that are fit poorly in a given model are given more weight in the next model [@elith2008]. Random forests and GBMs can provide strong predictive performance and fit highly non-linear relationships [@elith2008; @hastie2009]. We fit random forest models with the `randomForest` package [@liaw2002] for \textsf{R} with the default argument values. We fit boosted regression tree models with the `gbm` package [@ridgeway2015] for \textsf{R}. Based on cross-validation with the **caret** \textsf{R} package, we fit GBMs with 2000 trees, an interaction depth of $6$, a learning rate (shrinkage parameter) of $0.01$, and all other arguments at their default values.

@@

Applied to the simulated dataset of known status, the individual models had variable success at recovering the mean (status) and slope (trajectory) of $B/B_\mathrm{MSY}$ in the last five years. All models exhibited a high degree of scatter around the one-to-one line of perfect status prediction (Fig. \ref{hexagon}). CMSY exhibited bimodal predictions of status (Fig. \ref{hexagon}a), but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a). COM-SIR and SSCOM both correctly identified a number of low status stocks, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). There was generally little correlation between true and predicted trajectory for any of the individual models (rank-order correlation = `r range_slope_corr_non_sscom`) besides SSCOM (correlation = `r slope_corr_sscom`; Figs \ref{scatter-sim-slope}a--d).

@@

When applied to the stock assessment database, the superensemble models --- trained exclusively on the simulated dataset --- generally performed as well or better than the best individual models. The mean, random forest, and GBM ensembles outperformed the mPRM method which is trained directly on the RAM Legacy Stock Assessment database itself (Fig. \ref{performance}b, Fig. \ref{hexagon-ram}). Compared to the individual models, the machine learning superensembles increased accuracy by `r mean_ram$mach_vs_ind_mare_fold`%, improved correlation from `r mean_ram$ind_corr_range` to `r mean_ram$mach_corr_range`, and reduced bias from `r mean_ram$ind_mre_range` to `r mean_ram$mach_mre_range`.

@@

Ensemble methods provide a useful approach to situations where conservation and environmental resource management decisions must be made on the basis of multiple, potentially contrasting estimates of status --- a situation common to many settings beyond fisheries. Compared to individual models of fish population status, ensemble methods were consistently the best or among the best across three performance dimensions (accuracy, bias, and rank-order correlation), two response variables (status and trajectory), two datasets (simulated and global fisheries), and multiple ensemble methods (from a simple average to machine learning superensembles). Our results suggest choosing a superensemble model that allows for non-parametric relationships, such as machine learning methods. These methods provide added insight into individual model behaviour and generally performed the best; however, even a simple average of predictions across multiple models may provide a more useful measure of the desired parameter than a single model in many cases.

@@

Combining predictions from multiple models via superensemble methods is also broadly useful in other subfields of ecology and fisheries science. Predictions about extinction risk are widely used at the national (e.g. the US Endangered Species Act and the Canadian Species at Risk Act) and international [e.g. the IUCN Red List, @iucn2015] levels. These risk assessments generally involve fitting regression models to outcomes for individual species along with predictors of species risk [e.g. @anderson2011a; @pinsky2011], or population dynamics models to data for individual species [e.g. @dfo2010]. Both types of models are prone to error caused by model-misspecification and therefore results are sensitive to decisions about model structure [@brooks2015]. Although there are options to account for potential model-misspecification in determination of species risk [e.g. coefficient averaging, @burnham2002; generalized modeling, @yeakel2011; or semi-parametric methods, @thorson2014], ensemble methods are a relatively simple way to combine predictions in a transparent manner. Beyond estimates of status and trajectory, ensemble methods could be used to increase the robustness of spatial predictions when designing networks of protected areas [@rassweiler2014], to forecast potential spatial shifts in species distribution given climate impacts [@harsch2014], or to evaluate the effect of conservation or environmental management actions when experimental or observational data are limited *(Walsh et al. 2012)*. <!-- insert reference here: Walsh, J.C., Wilson, K.A., Benshemesh, J., Possingham, H.P., 2012. Unexpected outcomes of invasive predator control: the importance of evaluating conservation management actions. Animal Conservation 15, 319–328. -->
